### 2018-05-26 Paper/intro.tex
intro, first pass ...

%\begin{enumerate}
%\item Need for efficient Surrogates - especially for complex models to enable forward UQ, 
%sensitivity studies, calibration, and experimental design (Include citations).
%\item Computational Hurdles: Polynomial Chaos and GP can be computationally 
%intractable and suffer from the curse of dimensionality (Include plots for PCE
%to motivate dimension reduction with citations for both PCE and GP).
%\item Sensitivity analyis - a potent tool for dimension reduction. However, SA can be
%computationally prohibitive. In fact, there are studies demonstrating the use of 
%surrogates to reduce costs associated with SA (include citations). 
%\item DGSM - brief introduction and citations. 
%\item Key contributions of the paper: 1) Methodology that exploits DGSM to reduce the
%dimensionality of the problem and thus enables efficient construction of surrogates.
%2) Application of the proposed methodology to investigate relative importance of
%parameters in the stillinger-weber potential, commonly used for studying phonon transport in silicon. Further, construct a reasonably accurate PC surrogate in the reduced 
%space and demonstrate computational advantage of this approach. .
%\end{enumerate}

The emerging field of uncertainty quantification (UQ) aims at methodologies for 
incorporating, characterizing, quantifying, propagating, and reducing the 
uncertainties associated with predictive models and simulations. For situations
involving complex physical models and compute-intensive simulations, an
efficient approach to construction of model surrogates is sought to enable UQ
in a tractable manner. Polynomial chaos expansion 
(PCE)~\cite{Xiu:2002,Ghanem:2003,Olivier:2010} and 
Gaussian Process (GP)~\cite{Rasmussen:2004} or Kriging~\cite{Stein:2012} are
among the most commonly used surrogates for scientific applications. However,
both approaches quickly become prohibitive when the set of uncertain model
inputs and parameters is large-dimensional. Specifically, in the case of a PCE,
a $d$-dimensional polynomial basis with a total order of truncation, $p$ 
requires model realizations at $(p+1)^d$ quadrature nodes to exactly determine the PC
coefficients using a fully-tensorized Gauss quadrature. Consider a case where $d$ = 8 which
could be a regarded as a low-dimensional problem, and $p$ = 5, requires model evaluations at
roughly 1.7 million quadrature nodes. However, the degree of exactness in numerical estimation 
of the coefficients as made possible with a fully-tensored quadrature might not be desired. In that
case, computational effort can be reduced considerably by using nested
quadratures~\cite{Gentleman:1972a,Gentleman:1972b,Novak:1999,Waldvogel:2006} which allow 
model realizations at a given level to be used at higher levels of accuracy as well as sparse grids
based on Smolyak tensorization~\cite{Smolyak:1963}. Additonally, sparse basis 
techniques~\cite{Peng:2014,Hampton:2015,Blatman:2011} have been used to reduce  the required
computational effort for constructing a reasonably accurate PC surrogate.

One of the primary objectives for constructing a model surrogate is to be able to perform 
parametric sensitivity analysis tractably. Sensitivity analysis provides an insight into the
relative contributions of the uncertain model parameters and inputs to the uncertainty in
predictions. The analysis could potentially reduce the dimensionality of the problem.
However, as discussed earlier, constructing the surrogate might still require a large amount
of computational resources. It can thus be understood that dimension reduction prior to
constructing the surrogate might lead to enormous computational savings. In this work,
we present a strategy for identifying and screening uncertain model parameters that 
are significantly less important that the rest, thereby reducing the dimensionality of the
problem and enabling the construction of a reduced-order surrogate (ROS). As discussed in
section~\ref{sec:method}, parameter ranks are determined using a \textit{screening-metric}
based on derivative-based global sensitivity 
measures~\cite{Sobol:2009,Sobol:2010,Lamboni:2013,Kucherenko:2009,Kucherenko:2016}. 
The strategy is agnostic to the choice of methodology for constructing the surrogate. However, 
we will be using PCE to demonstrate the suitability of the proposed strategy. Below, in 
section~\ref{sec:bg}, we provide a brief introduction to derivative-based global sensitivity analysis
in the present context as well as the polynomial chaos methodology. Later, in section~\ref{sec:examples},
we motivate the proposed strategy by implementing it to a suite of simple applications, namely, the 
classic borehole function, a non-linear oscillator, and a semi-linear elliptic PDE. Finally, we demonstrate
suitability of our approach by using it to reduce the dimensionality of a H$_2$/O$_2$ kinetics problem
involving 19 reactions in section~\ref{sec:app}.

>>>>

The emerging field of uncertainty quantification (UQ) aims at methodologies for
incorporating, characterizing, quantifying, propagating, and reducing the
uncertainties associated with predictive models and simulations.  For
situations involving complex physical models and computationally intensive
simulations surrogate modeling often provides orders of magnitude speedups in
statistical studies. This is done by replacing repeated evaluations of
computationally expensive models by inexpensive evaluations of a surrogate
model.  Thus, an efficient approach to construction of surrogate models is of
central importance in enabling efficient uncertainty quantification for
computationally intensive models.

Commonly used surrogate modeling approaches use polynomial chaos expansions
(PCEs)~\cite{Xiu:2002,Ghanem:2003,Olivier:2010}, multivariate adaptive
regression splines (MARS)~\cite{friedman93}, Gaussian processes
(GPs)~\cite{Rasmussen:2004}, or kriging~\cite{Stein:2012}.  Many real-world
applications involve a large number of uncertain parameters. This makes
construction of surrogate models difficult or impossible in some cases.
However, in many situations, the variability in model observables of interest
is sensitive to only a small subset of the uncertain inputs.  Hence,
identifying model inputs that are inessential to variability in model output is
a key step that can facilitate surrogate model construction.

%However, both approaches
%quickly become prohibitive when the set of uncertain model inputs and
%parameters is large-dimensional. Specifically, in the case of a PCE, a
%$d$-dimensional polynomial basis with a total order of truncation, $p$ requires
%model realizations at $(p+1)^d$ quadrature nodes to exactly determine the PC
%coefficients using a fully-tensorized Gauss quadrature. Consider a case where
%$d$ = 8 which could be a regarded as a low-dimensional problem, and $p$ = 5,
%requires model evaluations at roughly 1.7 million quadrature nodes. However,
%the degree of exactness in numerical estimation of the coefficients as made
%%possible with a fully-tensored quadrature might not be desired. In that case,
%computational effort can be reduced considerably by using nested
%quadratures~\cite{Gentleman:1972a,Gentleman:1972b,Novak:1999,Waldvogel:2006}
%%which allow model realizations at a given level to be used at higher levels of
%accuracy as well as sparse grids based on Smolyak
%tensorization~\cite{Smolyak:1963}. Additionally, sparse basis
%techniques~\cite{Peng:2014,Hampton:2015,Blatman:2011} have been used to reduce
%the required computational effort for constructing a reasonably accurate PC
%surrogate.

%One of the primary objectives for constructing a model surrogate is to be able
%to perform parametric sensitivity analysis tractably.

Variance based global sensitivity analysis based on Sobol indices~\cite{Sobol93,
Sobol:2001,
Owen13,
SaltelliRattoAndresEtAl08} provides
insight into the relative contributions of the uncertain model inputs to the
uncertainty in predictions. Specifically, such analysis can help reduce the
dimensionality of the problem.  Computing Sobol indices, however, is a
computationally challenging task.  
Availability of a surrogate model typically
enables efficient computation 
of Sobol indices~\cite{Sudret08,CrestauxLeMaitreMartinez09,BlatmanSudret10,HartAlexanderianGremaud17,Sargsyan17}. This has enabled performing global sensitivity analysis
on a wide range of applications including in ocean 
modeling~\cite{AlexanderianWinokurSrajEtAl12,LiIskandaraniLeHenaffEtAl16},
geosciences~\cite{Namhata2016OladyshkinDilmoreEtAl16,deman2016,SaadAlexanderianPrudhommeEtAl17},
and chemical kinetics~\cite{DegasperiGilmore08,navarro2016global}, to name a few.

While surrogate models provide an efficient way of computing sensitivity
indices, constructing them in the case of models with high-dimensional inputs
can be as expensive as computing the Sobol indices via sampling.  Hence,
dimension reduction prior surrogate model construction can lead to enormous
computational savings.  Derivative based global sensitivity
analysis~\cite{Sobol:2009,Sobol:2010,Lamboni:2013,Kucherenko:2009,Kucherenko:2016}
often lead to a more tractable approach for global sensitivity
analysis~\cite{Kucherenko:2016}.  Also, the links between derivative based
global sensitivity measures (DGSMs) and total Sobol
indices~\cite{Sobol:2009,Kucherenko:2009,Kucherenko:2016} provide a strong
basis for their use in identifying unimportant parameters.

%However, as discussed
%earlier, constructing the surrogate might still require a large amount of
%computational resources.

\paragraph{Our approach}
We present a strategy for identifying and screening uncertain model parameters
that are significantly less important that the rest, thereby reducing the
dimensionality of the problem and enabling the construction of a reduced-order
surrogate (ROS).  Our approach combines DGSMs and surrogate modeling in an
iterative manner.  To make maximum use of computational resources, batches of
model evaluations are performed iteratively, and convergence of our DGSM based
screening metric is tested successively. Moreover, the validation steps built
in our method enable monitoring the accuracy of the parameter-screening and the
constructed surrogate model.  Our approach is agnostic to the choice of
methodology for constructing the surrogate. However, in the present work, we
rely on sparse PCEs to demonstrate the suitability of the proposed strategy.

%Moreover, the availability of adjoint solvers for
%efficient gradient computation can enhance the performance of our approach.

\paragraph{Contributions}
The contributions of this article are as follows: (i) we establish a robust and
practical framework for dimension reduction and surrogate modeling.  Our
approach is general in that it is applicable to a wide range of applications.
(ii) We present comprehensive numerical results demonstrating the viability of
our strategy in a number of motivating applications: the classic borehole
function, a nonlinear oscillator, and a semilinear elliptic PDE.  (iii) We
deploy our strategy in a application problem from chemical kinetics with 19
uncertain parameters. The application problem is studied in multiple regimes.
It is shown that the 19-parameter problem can be efficiently reduced to a 3- 
or 4-dimensional parameter problem.

\paragraph{Paper outline}
This article is structured as follows. In section~\ref{sec:bg}, we provide a
brief introduction to DGSMs as well as the polynomial chaos methodology used in
the present work.  In Section~\ref{sec:method} we present our proposed
approach. There we also provide a detailed numerical algorithm and a flow
diagram to aid practitioners in implementing the presented framework.
Section~\ref{sec:examples} is devoted to a number of numerical examples
examining various aspects of our approach. This is followed by implementation
of our framework in an H$_2$/O$_2$ chemical kinetics, in section~\ref{sec:app}.
Finally, concluding remarks are provided in section~\ref{sec:disc}.
